Electrifying a Bus Network

### 2022-12-14 project/Paper.Rmd
update paper

library(tidyverse)
library(ggthemes)
```

@@

The UVX is a bus rapid transit service operated by UTA and running through Provo and Orem, Utah. The service uses New Flyer XDE 60 model vehicles, so to make as close of a comparison as possible, I used the New Flyer XE 60 as the electric bus model for my calculations. This model is nearly identical to the XDE 60, with the main difference being that the XE is BE powered and the XDE is diesel-electric <><><>. Table XXX lists information about the current UVX route, and Table XXX lists details of the XE 60. Note that the current UVX fleet size is estimated based on vehicle tracking services, and several numbers are estimated from scheduling information.

TABLE

@@

A cost comparison of these charging solutions is given in Table XXX. Many of these values are estimated based on available information. It is worth noting that these costs can significantly vary depending on the specific implementation, but the numbers presented represent a rough average cost. Note also that the expected life span of this equipment (including the vehicles) is about 10\textendash 14 years.

TABLE (MO ebus and reg bus)

@@

An interesting result I immediately found is that given the specific scenario specifications, plug-in charging and overhead charging had nearly identical costs. Because of this, I averaged the cost of the two options and created a "Site Charging" scenario, as opposed to the "Inductive Lanes" scenario. Figures \@ref(fig:site-fig) and \@ref(fig:lanes-fig) show the cumulative cost over time of each BE bus adoption strategy. The shaded area represents the typical lifespan of the buses and charging equipment (10\textendash 14 years), so after this time the costs increase due to the need for equipment replacement. While in reality equipment would be replaced gradually rather than all at once, the cumulative effect on cost would be similar enough to not significantly affect the conclusions of this analysis.

```{r}
#| lanes-fig
#| fig.cap = "Cumulative cost based on site charging. The shaded area indicates the typical lifespan of equipment and vehicles, so replacement costs are added at this time."

@@

```{r}
#| lanes-fig
#| fig.cap = "Cumulative cost based on inductive lane charging. The shaded area indicates the typical lifespan of equipment and vehicles, so replacement costs are added at this time."

@@

Firstly, as inductive lanes are significantly more expensive than site chargers, it takes a longer time for any BE adoption strategy to save money compared to the baseline. In fact, no strategy is cheaper than the baseline before the estimated life span of the vehicles in the inductive lanes scenario. Though even the current diesel-electric buses would need to be replaced at that time, so ultimately the inductive lanes provide a reasonable solution after XXX years. The "Site Charging" scenario pays itself off much more quickly, especially in the 100% replacement strategy.

It is important to note that UTA already has a fleet of diesel-electric buses, so capital costs are only incurred for new BE buses and charging infrastructure. This is why the "Gradual" strategy increases in cost so quickly; new buses are being purchased each year, but a large portion of the fleet still has a high fuel cost per mile until most of the vehicles are replaced. Incidentally, after 11 years, when all vehicles have been replaced in the "Gradual" strategy, the difference between the "Gradual" and "100%" strategies is purely the difference in fuel costs, which is quite substantial.

- overhead/plugin about the same
- site charging > lanes (for UVX)
- economically viable

@@

Conclude that even with these limitations, transit agencies should look at BE buses. Also emissions.

>>>>

library(tidyverse)
library(ggthemes)
library(kableExtra)
```

@@

The UVX is a bus rapid transit service operated by UTA and running through Provo and Orem, Utah. The service uses New Flyer XDE 60 model vehicles, so to make as close of a comparison as possible, I used the New Flyer XE 60 as the electric bus model for my calculations. This model is nearly identical to the XDE 60, with the main difference being that the XE is BE powered and the XDE is diesel-electric <><><>. Table \@ref(tab:uvx) lists information about the current UVX route, and Table \@ref(tab:xe) lists details of the XE 60. Note that the current UVX fleet size is estimated based on vehicle tracking services, and several numbers are estimated from scheduling information.

```{r}
#| uvx

uvx <- read_csv("data/uvx.csv", col_names = FALSE)

uvx %>% 
	kbl(
		align = 'lrl',
		col.names = c("Utah Valley Express (830X)", "", ""),
		booktabs = TRUE,
		caption = "UVX Route Information") %>% 
	kable_styling(latex_options = "hold_position")
```

```{r}
#| xe

xe <- read_csv("data/xe60.csv", col_names = FALSE)

xe %>% 
	kbl(
		align = 'lrl',
		col.names = c("New Flyer XE 60", "", ""),
		booktabs = TRUE,
		caption = "New Flyer XE 60 Information") %>% 
	kable_styling(latex_options = "hold_position")
```

@@

A cost comparison of these charging solutions is given in Table \@ref(tab:costs). Many of these values are estimated based on available information. It is worth noting that these costs can significantly vary depending on the specific implementation, but the numbers presented represent a rough average cost. Note also that the expected life span of this equipment (including the vehicles) is about 10\textendash 14 years.

```{r}
#| costs

tribble(
	~item, ~cap, ~mo,
	"XE 60 Bus", 750000, 7000,
	"XDE 60 Bus", 1000000, 50000,
	"Plug-in Charger", 20000, 400,
	"Overhead Charger", 40000, 800,
	"Inductive Lanes (per mile)", 322000, 6400
) %>% 
	kbl(
		col.names = c("", "Capital Cost (USD)", "Maintenance Cost (USD/yr)"),
		caption = "Fleet Costs",
		booktabs = TRUE
	) %>% 
	kable_styling(latex_options = "hold_position")
```

@@

An interesting result I immediately found is that given the specific scenario specifications, plug-in charging and overhead charging had nearly identical costs. Because of this, I averaged the cost of the two options and created a "Site Charging" scenario, as opposed to the "Inductive Lanes" scenario. Figures \@ref(fig:site) and \@ref(fig:lanes) show the cumulative cost over time of each BE bus adoption strategy. The shaded area represents the typical lifespan of the buses and charging equipment (10\textendash 14 years), so after this time the costs increase due to the need for equipment replacement. While in reality equipment would be replaced gradually rather than all at once, the cumulative effect on cost would be similar enough to not significantly affect the conclusions of this analysis.

```{r}
#| site,
#| fig.height = 4,
#| fig.cap = "Cumulative cost based on site charging. The shaded area indicates the typical lifespan of equipment and vehicles, so replacement costs are added at this time.",
#| fig.pos = "h"

@@

```{r}
#| lanes,
#| fig.height = 4,
#| fig.cap = "Cumulative cost based on inductive lane charging. The shaded area indicates the typical lifespan of equipment and vehicles, so replacement costs are added at this time.",
#| fig.pos = "h"

@@

Firstly, as inductive lanes are significantly more expensive than site chargers, it takes a longer time for any BE adoption strategy to save money compared to the baseline. In fact, no strategy is cheaper than the baseline before the estimated life span of the vehicles in the inductive lanes scenario. Though even the current diesel-electric buses would need to be replaced at that time, so ultimately the inductive lanes provide a reasonable solution after 15\textendash 20 years. The "Site Charging" scenario pays itself off much more quickly (11\textendash 13 years), especially in the 100% replacement strategy.

It is important to note that UTA already has a fleet of diesel-electric buses, so capital costs are only incurred for new BE buses and charging infrastructure. This is why the "Gradual" strategy increases in cost so quickly; new buses are being purchased each year, but a large portion of the fleet still has a high fuel cost per mile and maintenance cost until most of the vehicles are replaced.

@@

Conclude that even with these limitations, transit agencies should look at BE buses. Also emissions.
- site charging > lanes (for UVX)
- economically viable

\newpage
